% main.tex
% Journal: Optimization Methods and Software (Taylor & Francis)
% Class:   interact.cls v1.05 (August 2017)
% Citation style: NUMERIC [N] via natbib  -- see PUBLICATION_GUIDELINES.md
% Last revised: March 2026

\documentclass[]{interact}  

% -------------------------------------------------------------------------
% SAFE PACKAGES (T&F production pipeline approved)
% -------------------------------------------------------------------------
\usepackage{epstopdf}                  % EPS figures with PDFLaTeX
\usepackage[caption=false]{subfig}     % sub-figures

% -------------------------------------------------------------------------
% TABLES AND FORMATTING
% -------------------------------------------------------------------------
\usepackage{booktabs}                  % \toprule, \midrule, \bottomrule
\usepackage{multirow}                  % \multirow{} for merged table rows
\usepackage{rotating}                  % sidewaystable for wide tables
\usepackage{xcolor}                    % for \textcolor{} in text annotations
\usepackage[section]{placeins}         % \FloatBarrier; [section] keeps floats in sections
\usepackage{float}                     % [H] placement for algorithms

% -------------------------------------------------------------------------
% DOUBLE-SPACING FOR PEER REVIEW (remove/comment out for camera-ready)
% -------------------------------------------------------------------------
\usepackage[doublespacing]{setspace}
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

% Uncomment both lines below to add line numbers for reviewers:
% \usepackage{lineno}

% -------------------------------------------------------------------------
% CITATION STYLE: NUMERIC [N]  (OMS journal requirement)
% -------------------------------------------------------------------------
\usepackage[numbers,sort&compress]{natbib}
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}

% -------------------------------------------------------------------------
% ALGORITHMS
% -------------------------------------------------------------------------
\usepackage{algorithm}
\usepackage{algpseudocode}

% -------------------------------------------------------------------------
% URLS IN REFERENCES
% -------------------------------------------------------------------------
\usepackage{url}

% -------------------------------------------------------------------------
% THEOREM-LIKE ENVIRONMENTS
% Theorem/Lemma/etc. share one counter scoped to sections (Theorem 2.1, Lemma 2.2, etc.)
% Remarks are numbered within sections (Remark 3.1, Remark 3.2, etc.)
% -------------------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{assumption}[theorem]{Assumption}

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{notation}{Notation}


% Starting the Paper

\begin{document}

\title{MPECSS: Scholtes Regularization with Adaptive Paths for Targeting B-stationary Points in MPECs}

\author{
\name{Saurabh (ORCID: 0009-0003-0458-8227)\textsuperscript{a}\thanks{CONTACT Saurabh.
      Email: 27098@arsd.du.ac.in}
\and
Kunwar Vijay Kumar Singh\textsuperscript{b}\thanks{CONTACT K.~V.~K.~Singh.
      Email: kunwar@du.ac.in}}
\affil{\textsuperscript{a}Department of Mathematics, Atma Ram Sanatan Dharma College, University of Delhi, India}
\affil{\textsuperscript{b}Department of Mathematics, Atma Ram Sanatan Dharma College, University of Delhi, India}
}

\maketitle


% ABSTRACT

\begin{abstract}
We propose MPECSS (MPEC Stationarity Search), an open-source algorithm for solving Mathematical Programs with Equilibrium Constraints (MPECs). Standard Scholtes regularization converges only to C-stationary points in general, a comparatively weak optimality concept providing no systematic mechanism for certifying the stronger B-stationary solutions required for genuine local optimality. MPECSS is implemented in Python via CasADi and IPOPT, with no commercial dependencies, fully open-source and reproducible. Convergence analysis establishes C-stationarity under standard assumptions; furthermore, MPECSS systematically certifies B-stationarity either directly when the multiplier sign test confirms S-stationarity under MPEC-LICQ or alternatively through LPEC (Linear Program with Equilibrium Constraints) verification. MPECSS achieves convergence rates of 96.3\% [87.5\%  B-stationary among converged], 92.1\% [85.4\% B-stationary among converged] and 80\% [85.4\% B-stationary among converged] on MacMPEC, MPECLib and NOSBENCH respectively.
\end{abstract}

\begin{keywords}
Complementarity constraints; Scholtes regularization; MPEC; Adaptive regularization; Open-source optimization
\end{keywords}

\begin{amscode}
90C33; 90C30; 65K05; 49M37; 90C90
\end{amscode}





% =========================================================================
% Section 1 -- Introduction
% =========================================================================
\input{section_introduction}

% =========================================================================
% Section 2 -- Preliminaries
% =========================================================================
\input{section_preliminaries}

% =========================================================================
% Section 3 -- The Proposed Algorithm: MPECSS
% =========================================================================
\input{section_algorithm}

% =========================================================================
% Section 4 -- Convergence Analysis
% =========================================================================
\input{section_convergence}

% =========================================================================
% Section 5 -- Numerical Experiments
% =========================================================================
\input{section_numerical}

% =========================================================================
% Section 6 -- Conclusions
% =========================================================================
\input{section_conclusions}

% =========================================================================
% Back-matter
% =========================================================================

\section*{Acknowledgements}

The authors thank the developers of CasADi and IPOPT for providing powerful open-source tools that enabled this research and the broader optimization community. The authors thank Anton Pozharskiy for translating the MacMPEC test collection to the CasADi format used in the numerical experiments. First author Saurabh want to specifically ackwolege the mentorship and support of their supervisor Dr. Kunwar Vijay Kumar Signh for providing guidance and significant intellectual contributions to the development of the MPECSS algorithm and the design of the numerical experiments.


\section*{Disclosure statement}

No potential conflict of interest was reported by the author(s).


\section*{Generative AI and AI-assisted technologies in the writing process}

During the preparation of this work the authors used the following
AI-assisted tools: Claude Opus 4 (Anthropic) and ChatGPT-5 (OpenAI) for brainstorming, deep
technical writing assistance, and exploration of algorithmic ideas;
Claude Sonnet 4 (Anthropic) for drafting and editing manuscript
sections; Claude Haiku 4.5 (Anthropic) for literature review and
language improvement; Gemini 3.1 Pro (Preview) (Google) for additional
literature review and language checking; and GitHub Copilot (Microsoft,
powered by OpenAI) for code development assistance. After using these
tools, the authors reviewed and edited all generated content and take
full responsibility for all aspects of this publication, including the
accuracy of mathematical derivations, the authenticity of reported
results, and compliance with publication ethics. These AI tools did not
make significant intellectual contributions to the research, nor did
they determine any aspect of the experimental outcomes.


\section*{Funding}

The authors report no funding.


\section*{Data availability statement}

The benchmark problem instances used in this study (MacMPEC, MPECLib,
and NOSBENCH) are publicly available as referenced in the text. The
checked-in repository contains the archived benchmark CSV files,
available run-environment metadata under
\texttt{results/outputs\_official\_*}, and the scripts used to generate
the reported figures and tables. These materials are available at
\url{https://github.com/mpecssalgorithm/MPECSS}.


\section*{Software availability}

The MPECSS algorithm is freely available under the Apache~2.0 licence.
The source code, benchmark datasets, and scripts required to reproduce
all reported results are available on GitHub at
\url{https://github.com/mpecssalgorithm/MPECSS}.
The archived official MacMPEC and MPECLib benchmark artifacts discussed
in this manuscript were produced at git commit
\texttt{33adefe} under Python~3.12.12 with CasADi~3.7.2,
NumPy~2.0.2, pandas~2.3.3, SciPy~1.16.3, psutil~5.9.5, and
matplotlib~3.10.0. The accompanying run-environment JSON records a
Linux execution environment on an Intel Xeon or AMD EPYC platform with
31.35\,GB RAM; the manuscript's numerical discussion refers to those
archived run-environment files where applicable. The repository also
contains a NOSBENCH official bundle at the same commit; because that
bundle is being refreshed after correcting duplicate timeout/crash CSV
rows, the NOSBENCH manuscript tables are marked as pending in the
current revision.


% =========================================================================
% References
% =========================================================================

\input{section_references}


\end{document}


