\section{Conclusions}
\label{sec:conclusions}

The central contribution is the integration of two complementary
mechanisms: an adaptive regularization-parameter update driven by the
implemented homotopy residual, and a multiplier sign test derived from
the theoretical characterizations of Scheel and
Scholtes~\cite{ScheelScholtes2000}. Together with the Phase~III
certification pipeline, these components give a practical mechanism for
searching for B-stationary points and certifying them when the final
LICQ/LPEC checks succeed.

The theoretical analysis established that MPECSS inherits the
convergence guarantees of Scholtes regularization, ensuring
C-stationarity under standard assumptions. More importantly, when the
internal multiplier sign test passes, the algorithm can promote the
iterate to a B-stationary solution through the subsequent Phase~III
LICQ/LPEC certification logic. Standard regularization methods do not
provide this implemented certification layer, and the manuscript now
distinguishes more carefully between the asymptotic Scholtes guarantee,
the finite-tolerance sign test, and the benchmark-harness
post-processing used in the reported experiments.

Comprehensive numerical experiments across three major benchmark collections demonstrated the practical effectiveness of the proposed approach. On the MacMPEC collection (191 problems), MPECSS achieved a convergence rate of 97.4\%, with 87.1\% of converged solutions certified as B-stationary; on MPECLib (88 problems, after excluding 4 non-standard MPEC formulations from the original 92), 88.6\% overall convergence with 83.3\% certified as B-stationary among converged solutions; and on NOSBENCH (603 problems), 92.0\% convergence with 84.9\% B-stationarity among converged solutions. These results highlight MPECSS as an open-source MPEC solver and benchmarking platform evaluated on all three benchmark suites in the present study. The implementation relies entirely on open-source software---Python, CasADi~\cite{AnderssonGillisHornRawlingsDiehl2019}, and IPOPT~\cite{WachterBiegler2006} with the MUMPS linear solver---eliminating dependencies on commercial tools such as MATLAB, SNOPT, Gurobi, or the HSL library.
Several directions remain open. On the theory side, the most important
next step is to close the remaining gap between the abstract
convergence statements and the exact implemented Phase~III
certification logic, especially in timeout- and branch-capped LPEC
cases. On the computational side, the large-scale \texttt{tinque},
\texttt{frictionalblock}, and traffic-network families suggest that the
next performance gains will come from stronger sparse linear-algebra
backends, tighter benchmark provenance, and more systematic
solver-to-solver comparison on curated archived result bundles.
