\begin{thebibliography}{99}

%1
\bibitem{AnderssonGillisHornRawlingsDiehl2019}
J.A.E. Andersson, J. Gillis, G. Horn, J.B. Rawlings, and M. Diehl,
\emph{CasADi: A software framework for nonlinear optimization and
optimal control}, Math. Program. Comput. 11(1) (2019), pp. 1--36.


%2
\bibitem{Anitescu2005}
M. Anitescu, \emph{On using the elastic mode in nonlinear programming
approaches to mathematical programs with complementarity constraints},
SIAM J. Optim. 15(4) (2005), pp. 1203--1236.


%3
\bibitem{Amestoy2001}
P.R. Amestoy, I.S. Duff, J.-Y. L'Excellent, and J. Koster, \emph{A fully asynchronous multifrontal solver using distributed dynamic scheduling}, SIAM J. Matrix Anal. Appl. 23(1) (2001), pp. 15--41.


%4
\bibitem{BaumruckerBiegler2008}
B.T. Baumrucker, J.G. Renfro, and L.T. Biegler, \emph{{MPEC} problem formulations and solution strategies with chemical engineering
applications}, Comput. Chem. Eng. 32(12) (2008), pp. 2903--2913.


%5
\bibitem{ColsonMarcotteSavard2007}
B. Colson, P. Marcotte, and G. Savard, \emph{An overview of bilevel optimization}, Ann. Oper. Res. 153(1) (2007), pp. 235--256.


%6
\bibitem{ConnGouldToint2000}
A.R. Conn, N.I.M. Gould, and Ph.L. Toint, \emph{Trust-Region Methods},
MPS-SIAM Series on Optimization, SIAM, Philadelphia, 2000.


%7
\bibitem{DeMiguel2005}
V. DeMiguel, M.P. Friedlander, F.J. Nogales, and S. Scholtes, \emph{A
two-sided relaxation scheme for mathematical programs with equilibrium
constraints}, SIAM J. Optim. 16(2) (2005), pp. 587--609.


%8
\bibitem{Dempe2002}
S. Dempe, \emph{Foundations of Bilevel Programming}, Kluwer Academic
Publishers, Dordrecht, 2002.


%9
\bibitem{MPECLib}
S. Dirkse and M. Bussieck, \emph{{MPECLib}: A collection of
mathematical programs with equilibrium constraints}, GAMS World test
library, available via \url{https://www.gams.com/modlib/libhtml/index.html}
and mirrored at \url{https://github.com/GAMS-dev/gamsworld}
(accessed March 2026).


%10
\bibitem{DolanMore2002}
E.D. Dolan and J.J. Mor\'{e}, \emph{Benchmarking optimization software
with performance profiles}, Math. Program. 91(2) (2002), pp. 201--213.


%11
\bibitem{FacchineiPang2003}
F. Facchinei and J.-S. Pang, \emph{Finite-Dimensional Variational
Inequalities and Complementarity Problems}, Springer Series in
Operations Research, Springer, New York, 2003.


%12
\bibitem{Ferreau2014qpOASES}
H.J. Ferreau, C. Kirches, A. Potschka, H.G. Bock, and M. Diehl,
\emph{{qpOASES}: A parametric active-set algorithm for quadratic
programming}, Math. Program. Comput. 6(4) (2014), pp. 327--363.


%13
\bibitem{FerrisPang1997}
M.C. Ferris and J.-S. Pang, \emph{Engineering and economic applications of complementarity problems},
SIAM Rev. 39(4) (1997), pp. 669--713.


%14
\bibitem{Fischer1992}
A. Fischer, \emph{A special Newton-type optimization method},
Optimization 24(3--4) (1992), pp. 269--284.


%15
\bibitem{FlegelKanzow2005}
M.L. Flegel and C. Kanzow, \emph{On M-stationary points for
mathematical programs with equilibrium constraints},
J. Math. Anal. Appl. 310(1) (2005), pp. 286--302.


%16
\bibitem{FletcherLeyffer2004}
R. Fletcher and S. Leyffer, \emph{Solving mathematical programs with
complementarity constraints as nonlinear programs}, Optim. Methods
Softw. 19(1) (2004), pp. 15--40.


%17
\bibitem{FletcherLeyfferRalphScholtes2006}
R. Fletcher, S. Leyffer, D. Ralph, and S. Scholtes, \emph{Local
convergence of {SQP} methods for mathematical programs with equilibrium
constraints}, SIAM J. Optim. 17(1) (2006), pp. 259--286.


%18
\bibitem{GillMurraySaunders2005SNOPT}
P.E. Gill, W. Murray, and M.A. Saunders, \emph{SNOPT: An SQP algorithm
for large-scale constrained optimization}, SIAM Rev. 47(1) (2005),
pp. 99--131.


%19
\bibitem{HoheiselKanzowSchwartz2013}
T. Hoheisel, C. Kanzow, and A. Schwartz, \emph{Theoretical and
numerical comparison of relaxation methods for mathematical programs
with complementarity constraints}, Math. Program. 137(1--2) (2013),
pp. 257--288.


%20
\bibitem{HuRalph2004}
X.M. Hu and D. Ralph, \emph{Convergence of a penalty method for
mathematical programming with complementarity constraints},
J. Optim. Theory Appl. 123(2) (2004), pp. 365--390.


%21
\bibitem{HuRalph2007}
X. Hu and D. Ralph, \emph{Using EPECs to model bilevel games in
restructured electricity markets with locational prices},
Oper. Res. 55(4) (2007), pp. 809--824.


%22
\bibitem{HuangfuHall2018}
Q. Huangfu and J.A.J. Hall, \emph{Parallelizing the dual revised simplex
method}, Math. Program. Comput. 10(1) (2018), pp. 119--142.


%23
\bibitem{Kadrani2009}
A. Kadrani, J.P. Dussault, and A. Benchakroun, \emph{A new
regularization scheme for mathematical programs with complementarity
constraints}, SIAM J. Optim. 20(1) (2009), pp. 78--103.


%24
\bibitem{KanzowSchwartz2013}
C. Kanzow and A. Schwartz, \emph{A new regularization method for
mathematical programs with complementarity constraints with strong
convergence properties}, SIAM J. Optim. 23(2) (2013), pp. 770--798.


%25
\bibitem{KanzowSchwartz2015}
C. Kanzow and A. Schwartz, \emph{The price of inexactness: convergence
properties of relaxation methods for mathematical programs with
complementarity constraints revisited}, Math. Oper. Res. 40(2) (2015),
pp. 253--275.


%26
\bibitem{Kanzow1996}
C. Kanzow, \emph{Some noninterior continuation methods for linear
complementarity problems}, SIAM J. Matrix Anal. Appl. 17(4) (1996),
pp. 851--868.


%27
\bibitem{KlarbringBjorkman1988}
A. Klarbring and G. Bj\"{o}rkman, \emph{A mathematical programming approach to contact problems with friction and varying contact surface},
Comput. Struct. 30(5) (1988), pp. 1185--1198.


%28
\bibitem{LeyfferLopezCalvaNocedal2006}
S. Leyffer, G. L\'{o}pez-Calva, and J. Nocedal, \emph{Interior methods
for mathematical programs with complementarity constraints},
SIAM J. Optim. 17(1) (2006), pp. 52--77.


%29
\bibitem{Leyffer2000}
S. Leyffer, \emph{MacMPEC: A collection of mathematical programs with
equilibrium constraints}, Technical Report ANL/MCS-P817-0500, Argonne
National Laboratory, 2000.


%30
\bibitem{LeyfferMunson2010}
S. Leyffer and T. Munson, \emph{Solving multi-leader-common-follower games},
Optim. Methods Softw. 25(4) (2010), pp. 601--623.


%31
\bibitem{LinFukushima2005}
G.-H. Lin and M. Fukushima, \emph{A modified relaxation scheme for
mathematical programs with complementarity constraints},
Ann. Oper. Res. 133(1--4) (2005), pp. 63--84.


%32
\bibitem{LuoPangRalph1996}
Z.-Q. Luo, J.-S. Pang, and D. Ralph, \emph{Mathematical Programs with
Equilibrium Constraints}, Cambridge University Press, Cambridge, 1996.


%33
\bibitem{MengYangBell2001}
Q. Meng, H. Yang, and M.G.H. Bell, \emph{An equivalent continuously differentiable model and a locally convergent algorithm for the continuous network design problem},
Transp. Res. Part B Methodol. 35(1) (2001), pp. 83--105.


%34
\bibitem{NocedalWright2006}
J. Nocedal and S.J. Wright, \emph{Numerical Optimization}, 2nd ed.,
Springer Series in Operations Research and Financial Engineering,
Springer, New York, 2006.


%35
\bibitem{NurkanovicLeyffer2025}
A. Nurkanovi\'{c} and S. Leyffer, \emph{A globally convergent method
for computing B-stationary points of mathematical programs with
equilibrium constraints}, arXiv:2501.13835v4, 2025.


%36
\bibitem{NurkanovicPozharskiyDiehl2024}
A. Nurkanovi\'{c}, A. Pozharskiy, and M. Diehl, \emph{Solving
mathematical programs with complementarity constraints arising in
nonsmooth optimal control}, Vietnam J. Math. (2024), pp. 1--39.


%37
\bibitem{OutrataKocvaraZowe1998}
J.V. Outrata, M. Ko\v{c}vara, and J. Zowe, \emph{Nonsmooth Approach to
Optimization Problems with Equilibrium Constraints: Theory, Applications
and Numerical Results}, Kluwer Academic Publishers, Dordrecht, 1998.


%38
\bibitem{RaghunnathanBiegler2005}
A.U. Raghunathan and L.T. Biegler, \emph{An interior point method for
mathematical programs with complementarity constraints (MPCCs)},
SIAM J. Optim. 15(3) (2005), pp. 720--750.


%39
\bibitem{RalphWright2004}
D. Ralph and S.J. Wright, \emph{Some properties of regularization and
penalization schemes for MPECs}, Optim. Methods Softw. 19(5) (2004),
pp. 527--556.


%40
\bibitem{RockafellarWets1998}
R.T. Rockafellar and R.J.-B. Wets, \emph{Variational Analysis},
Springer, Berlin, 1998.


%41
\bibitem{ScheelScholtes2000}
H. Scheel and S. Scholtes, \emph{Mathematical programs with
complementarity constraints: stationarity, optimality, and sensitivity},
Math. Oper. Res. 25(1) (2000), pp. 1--22.


%42
\bibitem{Scholtes2001}
S. Scholtes, \emph{Convergence properties of a regularization scheme
for mathematical programs with complementarity constraints},
SIAM J. Optim. 11(4) (2001), pp. 918--936.


%43
\bibitem{SteffensenUlbrich2010}
S. Steffensen and M. Ulbrich, \emph{A new relaxation scheme for
mathematical programs with equilibrium constraints},
SIAM J. Optim. 20(5) (2010), pp. 2504--2539.


%44
\bibitem{WachterBiegler2006}
A. W\"{a}chter and L.T. Biegler, \emph{On the implementation of an
interior-point filter line-search algorithm for large-scale nonlinear
programming}, Math. Program. 106(1) (2006), pp. 25--57.


%45
\bibitem{Ye2005}
J.J. Ye, \emph{Necessary and sufficient optimality conditions for
mathematical programs with equilibrium constraints},
J. Math. Anal. Appl. 307(1) (2005), pp. 350--369.


\end{thebibliography}
