\section{The Proposed Algorithm: MPECSS}
\label{sec:algorithm}
In this section, we present the MPECSS algorithm, which combines Scholtes's regularization \cite{Scholtes2001} with an adaptive path-following strategy guided by a multiplier sign test and complementarity residual to target B-stationary points of MPECs. We first describe the main parts of the algorithm and then present a detailed step-by-step procedure.

\subsection{Scholtes's regularization}
\label{sec:scholtes}
As discussed in Section~\ref{sec:intro}, the main source of difficulty in MPECs~\eqref{eq:general_mpec} is the complementarity conditions, which cause standard constraint qualifications to fail. In 2001, Scholtes~\cite{Scholtes2001} proposed a regularization technique, which relaxed the complementarity constraints $G_\ell(x) \cdot H_\ell(x) = 0$ to $G_\ell(x) \cdot H_\ell(x) \leq t_k$ for a sequence of positive parameters $t_k \to 0$ indexed by outer iteration $k$. This regularization transforms the original MPEC into a sequence of smooth NLP subproblems that can be solved using standard NLP solvers~\cite{FletcherLeyffer2004}. The key idea is that as $t_k$ approaches zero, the solutions of the regularized problems approach a stationary point of the original MPEC.

After replacing the complementarity constraints with their regularized counterparts, the regularized problem $\text{NLP}(t_k)$ with parameter $t_k$ at outer iteration $k$ can be formulated as~\cite{Scholtes2001}:
\begin{equation}
\begin{aligned}
  \text{NLP}(t_k): \quad & \min_{x \in \mathbb{R}^n} \quad && f(x) \\
  & \text{s.t.} \quad && g_i(x) \leq 0, \quad i = 1, \ldots, m, \\
  & && h_j(x) = 0, \quad j = 1, \ldots, p, \\
  & && G_\ell(x) \ge 0, \quad \ell = 1, \ldots, q, \\
  & && H_\ell(x) \ge 0, \quad \ell = 1, \ldots, q, \\
  & && G_\ell(x) \cdot H_\ell(x) \leq t_k, \quad \ell = 1, \ldots, q.
\end{aligned}
\label{eq:nlp}
\end{equation}

The code accepts benchmark models whose native complementarity format is
not exactly the textbook form~\eqref{eq:nlp}. For MPECLib and
NOSBENCH, finite lower bounds on the complementarity functions are
shifted into internal variables
\[
  \widehat G_\ell(x) := G_\ell(x)-\ell^G_\ell,\qquad
  \widehat H_\ell(x) := H_\ell(x)-\ell^H_\ell,
\]
so that the solver always works with nonnegative internal pairs
$\widehat G_\ell,\widehat H_\ell \ge 0$. For MacMPEC, the parser also
supports mixed and box-complementarity structures, including free
$G$-components and upper-bound pairs of the form
$(-G_\ell(x))(u_\ell-H_\ell(x)) \le t_k$. The notation in the sequel is
therefore to be understood as the solver's internal complementarity
representation after this benchmark-specific normalization.

The sequence of solutions $\{\text{NLP}(t_k)\}_{k=0}^\infty$ is expected to converge to a C-stationary point of the original MPEC as $t_k \to 0$~\cite{Scholtes2001}.

The regularization sequence $\{t_k\}$ is generated by a geometric reduction~\cite{Scholtes2001,HoheiselKanzowSchwartz2013}:
\begin{equation}
  t_{k+1} = \kappa \cdot t_k, \quad k = 0, 1, 2, \ldots,
  \label{eq:param_update}
\end{equation}
where $t_0 > 0$ is the initial regularization parameter and $\kappa \in (0, 1)$ is the reduction factor. As $k \to \infty$, $t_k \to 0$, bringing the regularized problems progressively closer to the original MPEC~\eqref{eq:general_mpec}. 

Each subproblem $\text{NLP}(t_{k+1})$ is initialized from the solution $x^k$ of the previous subproblem $\text{NLP}(t_k)$. This \emph{warm-start} strategy means the solver follows a continuous \emph{regularization path} $\{x^k\}$ as $t_k$ decreases, rather than solving each subproblem from scratch. This makes the method a path-following scheme~\cite{RalphWright2004,HoheiselKanzowSchwartz2013} and greatly speeds up computation.

A solution $x^k$ of $\text{NLP}(t_k)$ is accepted when the KKT conditions are satisfied to a tolerance $\varepsilon_k > 0$. In practice, $\varepsilon_k$ is reduced alongside $t_k$ (e.g., $\varepsilon_k = O(t_k)$; see~\cite{HoheiselKanzowSchwartz2013,KanzowSchwartz2015}) to ensure that the accuracy of the solutions improves as $t_k$ gets smaller.

Scholtes~\cite{Scholtes2001} proved that limit points of the regularized sequence are at best C-stationary for the MPEC. In general, C-stationarity is \emph{strictly weaker} than B-stationary
 and the standard method provides no mechanism to detect or target B-stationary points~\cite{Scholtes2001,HoheiselKanzowSchwartz2013}. Alternative relaxation formulations have been proposed that guarantee stronger stationarity under weaker
 conditions (e.g., Kanzow and Schwartz~\cite{KanzowSchwartz2013} achieve
 M-stationarity under MPEC-LICQ; DeMiguel et al.~\cite{DeMiguel2005} propose a two-sided relaxation), but their theoretical stationarity guarantees can
 fail when NLP solvers return approximate solutions~\cite{KanzowSchwartz2015}. MPECSS
 takes a different approach: it preserves Scholtes's product relaxation which is
 numerically robust and fills the stationarity gap by combining the regularization path
 with an \emph{adaptive path-following strategy} guided by
the \emph{multiplier sign test} (Section~\ref{sec:stationarity}). At each
outer iteration, the sign test checks the $G$- and $H$-multipliers of
$\text{NLP}(t_k)$ to confirm (or rule out) S-stationarity of the
current iterate and uses this information to adjust $\{t_k\}$
accordingly. The full description of this adaptive mechanism is given
in Sections~3.3--3.4. Notably, S-stationarity implies B-stationarity by
Scheel and Scholtes~\citep{ScheelScholtes2000}, and under the
MPEC-LICQ constraint qualification, S-stationarity and B-stationarity
are equivalent. In the implementation, however, a sign-test pass is
carried into Phase~III, where B-stationarity is certified either by
LICQ equivalence or by the LPEC check when LICQ is absent or
inconclusive.

\subsection{Phase~I: Feasibility}
\label{sec:feasibility}

The main MPECSS loop (Section~\ref{sec:adaptive_strategy}) requires a starting point that approximately satisfies the complementarity conditions $G_\ell(x)^{\top}\, H_\ell(x) \approx 0$ with $G_\ell(x) \geq 0$ and $H_\ell(x) \geq 0$. A user-provided initial point~$x_0$ may be far from complementarity-feasible (i.e., the products $G_\ell(x_0)^{\top}\, H_\ell(x_0)$ may be large). To address this, MPECSS begins with an optional \emph{feasibility phase} (Phase~I), following standard MPEC feasibility preprocessing approaches~\cite{BaumruckerBiegler2008,NurkanovicLeyffer2025}, that finds a good starting point satisfying the complementarity conditions. Before solving, the initial point~$x_0$ is pushed into the strict interior of the variable bounds by a fraction $\nu = 0.1$, i.e., for each dimension~$i$ with finite bounds $[\ell_i, u_i]$, the starting value is clipped to $[\ell_i + \nu(u_i - \ell_i),\; u_i - \nu(u_i - \ell_i)]$. This \emph{interior push} avoids Jacobian singularities that interior-point methods can encounter when initialised exactly on variable bounds~\cite{WachterBiegler2006}.

Phase~I then solves a sequence of up to three progressively more robust NLP formulations to find a complementarity-feasible point. The general structure minimizes a violation measure $\Theta(x)$ of the complementarity conditions, subject to the original MPEC feasibility constraints:
\begin{equation}
\begin{aligned}
  & \min_{x \in \mathbb{R}^n} \quad && \Theta(x) \\
  & \text{s.t.} \quad && g_i(x) \leq 0, \quad i = 1, \ldots, m, \\
  & && h_j(x) = 0, \quad j = 1, \ldots, p, \\
  & && G_\ell(x) \geq 0, \quad \ell = 1, \ldots, q, \\
  & && H_\ell(x) \geq 0, \quad \ell = 1, \ldots, q.
\end{aligned}
\label{eq:phase_i}
\end{equation}
The formulation of the objective $\Theta(x)$ is progressively adjusted in a cascade if the previous attempt fails to achieve a satisfactory complementarity residual ($r < 10^{-6}$). These attempts correspond to: (1) \emph{L2 product minimisation} $\Theta(x) = \sum_{\ell=1}^q \big(G_\ell(x) H_\ell(x)\big)^2$ to standardly push products to zero; (2) \emph{Smooth L1 minimisation} $\Theta(x) = \sum_{\ell=1}^q \sqrt{\big(G_\ell(x) H_\ell(x)\big)^2 + \epsilon}$ (with $\epsilon = 10^{-10}$) to provide a less aggressive gradient when ill-conditioned; and (3) an \emph{Epigraph formulation} which solves $\min_{x, \tau} \tau$ subject to the constraints in~\eqref{eq:phase_i} as well as $\big(G_\ell(x) H_\ell(x)\big)^2 \leq \tau$ for all $\ell = 1, \ldots, q$ and $\tau \geq 0$, yielding stable gradients on severely infeasible problems. In cases involving box-constrained complementaries, the formulation naturally aggregates both the lower complementarity pair $G_\ell(x) H_\ell(x)$ and the upper pair $G_\ell(x)(u_\ell - H_\ell(x))$ if a finite upper bound $u_\ell$ exists.

No regularization parameter appears in any of these formulations; each is a standard smooth NLP solved by IPOPT~\cite{WachterBiegler2006}. Upon convergence of any formulation, the resulting point $\bar{x}$ satisfies $G_\ell(\bar{x}) \cdot H_\ell(\bar{x}) \approx 0$ for all~$\ell$ and lies in the feasible region of the original constraints. This point is then used as the warm-start~$x^0$ for the main Scholtes outer loop.

Three baseline safety mechanisms augment Phase~I's reliability. First,
a \textit{structured multistart} deploys deterministic bound/midpoint
candidates together with Gaussian-perturbed starts to improve robustness
when the user-supplied initial point is poor. Second, a
\textit{displacement guard} rejects candidates drifting excessively
($\|\bar{x}^{(s)} - x_0\|_R \geq D_{\max}$) from the user-provided
region to retain local relevance. Third, the implementation records the
Phase~I objective values for diagnostics, but the actual acceptance rule
is driven by improved complementarity residual together with the
displacement safeguard, rather than by a separate original-objective
filter.

In practice, the solver may fail to find an initial search direction from the user-provided starting point, immediately reporting infeasibility. To improve startup robustness without changing the theoretical Scholtes framework~\cite{Scholtes2001}, MPECSS cycles through the three progressively more robust objective formulations described above (L2, smooth~L1, epigraph). Each formulation provides a different gradient landscape, and the sequence is designed so that when one formulation encounters numerical difficulties, the next one often succeeds. For very large problems ($n > 3000$), Phase~I is skipped entirely to avoid excessive computational cost; in this case, the user-provided starting point is passed directly to Phase~II.

Algorithm~\ref{alg:phase1} summarizes the practical Phase~I procedure used to generate a reliable warm start for the adaptive outer loop.

\begin{algorithm}[t]
\caption{Phase~I Feasibility Bootstrap}
\label{alg:phase1}
\begin{algorithmic}[1]
\Require Initial point $x_0$, bounds $\ell,u$, feasibility threshold $\eta_f$, displacement limit $D_{\max}$, interior fraction $\nu$
\Ensure Warm-start point $x^0$ for Phase~II
\If{$n > 3000$}
  \State \Return $x^0 \gets x_0$ \Comment{Skip Phase~I for very large problems}
\EndIf
\State $x_0 \gets \textsc{InteriorPush}(x_0, \ell, u, \nu)$ \Comment{Push into strict interior of bounds}
\For{attempt $a = 0, 1, 2$ (L2, smooth L1, epigraph)}
  \State Solve Phase~I NLP with objective regime~$a$ from $x_0$ (or incumbent) and compute residual $r_a$
  \State Accept if $r_a < r_{\text{best}}$ and $\|\bar{x} - x_0\|/\max(1,\|x_0\|) < D_{\max}$
  \If{$r_a < \eta_f$}
    \State \Return $x^0 \gets \bar{x}$
  \EndIf
\EndFor
\State Build multistart set: lower-bound, upper-bound, midpoint corners, and Gaussian-perturbed candidates
\For{each candidate start $x^{(s)}$ in multistart set}
  \For{attempt $a = 0, 1$ (L2, smooth L1)}
    \State Solve Phase~I NLP with regime~$a$ from $x^{(s)}$ and obtain candidate $\bar{x}^{(s)}$
    \State Compute residual $r^{(s)}$ and normalised displacement $d^{(s)}$
    \If{$r^{(s)} < r_{\text{best}}$ and $d^{(s)} < D_{\max}$}
      \State Update incumbent $x^0 \gets \bar{x}^{(s)}$, $r_{\text{best}} \gets r^{(s)}$
    \EndIf
  \EndFor
  \If{$r_{\text{best}} < \eta_f$}
    \State \textbf{break}
  \EndIf
\EndFor
\State \Return $x^0$
\end{algorithmic}
\end{algorithm}

When Phase~I achieves a complementarity residual $r_0 \leq 100\,
\varepsilon_{\mathrm{tol}}$, the implementation performs an early
stationarity evaluation before entering the Phase~II outer loop: it
solves one NLP at a small regularization parameter
$\bar{t} = \max(0.1\,r_0,\, 10^{-12})$ to extract multipliers and runs
the sign test.  Three outcomes are distinguished:
If the sign test passes and $r_0 \leq \varepsilon_{\mathrm{tol}}$,
Phase~II is skipped and the point proceeds directly to Phase~III.
If $r_0 \leq \varepsilon_{\mathrm{tol}}$ but the sign test fails
and the biactive count satisfies
$|\mathcal{I}_{00}| \leq N_{\mathrm{LPEC}}$ (default $N_{\mathrm{LPEC}}=15$),
Phase~III is invoked directly since the LPEC enumeration
($2^{|\mathcal{I}_{00}|}$ branches) is tractable.
Otherwise, Phase~II runs normally.
Additionally, when Phase~I produces $r_0 < t_0$, the initial
regularization parameter is fast-forwarded to
$t_0 \gets \max(10\,r_0,\, \varepsilon_{\mathrm{tol}}\,\tau)$ to
preserve the refined starting point and avoid redundant outer
iterations at large $t$ values.


\subsection{Phase II: Adaptive Path-Following Strategy with Sign Test}
\label{sec:adaptive_strategy}
The main new idea in MPECSS is to integrate a \emph{multiplier sign test} 
into the regularization path-following framework to guide the sequence  
$\{t_k\}$ towards B-stationary points~\cite{ScheelScholtes2000}. The main idea is as follows:

Algorithm~\ref{alg:phase2-main} gives the top-level orchestration of Phase~II, while Algorithms~\ref{alg:phase2-update} and~\ref{alg:phase2-restoration} detail its two key subroutines.

\begin{algorithm}[t]
\caption{Phase~II Main Adaptive Path-Following Loop}
\label{alg:phase2-main}
\begin{algorithmic}[1]
\Require Warm-start $x^0$, initial regularization $t_0$, reduction factor $\kappa$, tolerance $\varepsilon_{\mathrm{tol}}$, maximum iterations $k_{\max}$
\Ensure Candidate point $x^*$ for Phase~III
\State Set $k \gets 0$; initialise history buffers, pathing/stagnation counters, and best-point pather $\mathit{best}$
\While{$k < k_{\max}$}
  \State Solve NLP$(t_k)$~\eqref{eq:nlp} from warm-start $x^k$ to obtain $(x^{k+1},\lambda^{k+1})$
  \State Convert multipliers via~\eqref{eq:multiplier_conversion}; compute homotopy residual $r_{k+1}$ via~\eqref{eq:comp_res}
  \State Run sign test on $\mathcal{I}_{00}(x^{k+1})$; update $\mathit{best}$ if $(r_{k+1},f_{k+1})$ improves the incumbent
  \If{$r_{k+1} \le \varepsilon_{\mathrm{tol}}$ \textbf{and} sign test passes}
    \State Set $x^* \gets x^{k+1}$ (complementarity-feasible S-stationarity candidate) and \textbf{break}
  \EndIf
  \If{sign test fails \textbf{and} $\mathcal{I}_{00}(x^{k+1}) \ne \emptyset$}
    \State Invoke restoration (Alg.~\ref{alg:phase2-restoration}) at fixed $t_k$; if it returns a point, replace $x^{k+1}$ and re-evaluate $r_{k+1}$ and the sign test
  \EndIf
  \State Update $t_{k+1}$ via Algorithm~\ref{alg:phase2-update}
  \State $k \gets k+1$
\EndWhile
\If{no convergence}
  \State $x^* \gets \mathit{best}$ (lowest complementarity residual; ties broken by objective value)
\EndIf
\State \Return $x^*$
\end{algorithmic}
\end{algorithm}

\subsubsection{Solve regularized NLP subproblems}
\label{sec:solveRegNLPs}
Phase~I provides a robust warm start for the main loop. At each outer
iteration $k$, MPECSS builds the smooth subproblem $\text{NLP}(t_k)$
in CasADi and solves it with IPOPT~\cite{WachterBiegler2006} as its
primary NLP engine. In the reported benchmark runs, IPOPT uses the
open-source MUMPS linear solver~\cite{Amestoy2001} with tuned settings
for robustness on ill-conditioned MPEC subproblems.\footnote{%
  MUMPS is configured with increased workspace allocation
  ($\texttt{mumps\_mem\_percent}=500$), Hungarian scaling
  ($\texttt{mumps\_scaling}=77$), and pivot tolerance
  $\texttt{mumps\_pivtol}=10^{-6}$. IPOPT uses gradient-based NLP
  scaling, up to 4 second-order correction steps
  ($\texttt{max\_soc}=4$), and 3 watchdog trial iterations.}
The code base also provides an optional small-problem
SQP$+$qpOASES~\cite{Ferreau2014qpOASES} route for problems with
$n \leq 400$ variables, when qpOASES is available. In the current
implementation and archived benchmark runs, interior-point methods are
used as the default because they were empirically the most reliable
option across the tested problem families.
For problems with $n \geq 500$, CasADi's MX symbolic type is used with reverse-mode automatic differentiation (\texttt{ad\_weight}$=1.0$), reducing Jacobian memory consumption from $O(n_x \times 64)$ to $O(n_g)$ compared to forward-mode; additionally, an LRU cache with automatic eviction is employed for solver templates and Jacobian functions to ensure stable memory usage during long benchmark campaigns.

When the primary solve fails (e.g., due to infeasibility detection,
restoration failure, or numerical errors), MPECSS applies an automatic
\emph{solver fallback chain}: (1)~retry with Mehrotra's
predictor-corrector algorithm and probing oracle;
(2)~retry with relaxed tolerances ($\mathrm{tol}=10^{-8}$,
$\mathrm{acceptable\_tol}=10^{-6}$);
(3)~retry with gradient-based NLP scaling and relaxed bound factor
($10^{-8}$); and (4)~if all IPOPT variants fail, switch to the
Fischer--Burmeister reformulation~\cite{Fischer1992,Kanzow1996} of the
complementarity constraints, which provides a different gradient
landscape and often succeeds where the product relaxation encounters
numerical difficulties.

Each subproblem $\text{NLP}(t_{k+1})$ is warm-started from the solution
$x^k$ of the previous subproblem $\text{NLP}(t_k)$, including primal
and dual variables.\footnote{%
  IPOPT warm-start parameters: $\texttt{warm\_start\_bound\_push} =
  \texttt{warm\_start\_bound\_frac} = \texttt{warm\_start\_slack\_bound\_frac}
  = \texttt{warm\_start\_slack\_bound\_push} = 10^{-9}$.}
This \emph{warm-start} strategy means the solver follows a continuous
\emph{regularization path} $\{x^k\}$ as $t_k$ decreases, rather than
solving each subproblem from scratch, greatly reducing the total number
of NLP iterations.


\subsubsection{Constraint stacking and multiplier sign rule}
\label{sec:stacking_signs}
For implementation consistency, the regularized constraints are passed to the NLP solver in a fixed order:
\begin{equation}
g(x)=\begin{bmatrix}
g_{\mathrm{orig}}(x)\\
G(x)+\delta_k\\
H(x)+\delta_k\\
G(x)\circ H(x)-t_k
\end{bmatrix},
\label{eq:stacked_constraints}
\end{equation}
where the tag $\mathrm{orig}$ means ``original'' and
$g_{\mathrm{orig}}(x)$ collects the inequality constraints
inherited from the original MPEC model (before adding regularization
terms), $\circ$ denotes componentwise product, and $\delta_k \geq 0$
is an optional regularization shift.\footnote{%
  In all reported numerical experiments, $\delta_k = 0$; the shift is
  retained in the formulation for generality.} Let
$\lambda^{\mathrm{cas}}$ be the solver multiplier vector associated
with~\eqref{eq:stacked_constraints}, where the tag ``$\mathrm{cas}$''
means multipliers returned in the CasADi/IPOPT solver convention.
Because CasADi returns non-positive multipliers at active lower bounds
for the $G$- and $H$-constraints, MPECSS converts these complementarity
constraints to MPEC-oriented multipliers by negation:
\begin{equation}
\gamma^{G,k} = -\lambda^{\mathrm{cas}}_{G},\qquad
\gamma^{H,k} = -\lambda^{\mathrm{cas}}_{H},
\label{eq:multiplier_conversion}
\end{equation}
The multiplier of the relaxed product constraint is still extracted and
stored for diagnostics, but it is \emph{not} part of the
S-stationarity sign test. All sign tests in MPECSS are therefore
performed only on the converted multipliers in
Equation~\eqref{eq:multiplier_conversion}.

\subsubsection{Multiplier Sign Test}
At each outer iteration $k$, after solving $\text{NLP}(t_k)$, we
extract the converted multipliers $(\gamma^{G,k},\gamma^{H,k})$ for
the shifted $G$- and $H$-constraints. For biactive indices
$\ell \in \mathcal{I}_{00}(x^k)$, the implemented sign test checks
\[
\gamma^{G,k}_\ell \ge -\tau,\qquad
\gamma^{H,k}_\ell \ge -\tau,
\]
with a tolerance chosen in the implementation as
$\tau=10^{-6}$ by default. The adaptive quantity
$\sigma_k=\max\{10^{-8},\sqrt{t_k}\}$ is used instead for detecting the
numerically biactive set. The Scholtes product multiplier is
not included in this test, because it is not part of the
S-stationarity definition. If the sign test passes and the homotopy
residual is sufficiently small, Phase~II returns an S-stationarity
candidate to Phase~III, where final B-stationarity is certified by the
implemented LICQ/LPEC logic.

\subsubsection{Adaptive $t$-update strategy}
\label{sec:adaptive_t}
In the standard Scholtes scheme, the regularization parameter is
reduced by a fixed geometric rule $t_{k+1} = \kappa \, t_k$~\cite{Scholtes2001}
regardless of the solver's progress. MPECSS replaces this with an
\emph{adaptive update strategy} that monitors the homotopy
complementarity residual
\begin{equation}
  r_k \;=\; \max_{\ell = 1,\ldots,q} \bigl|G_\ell(x^k)\,H_\ell(x^k)\bigr|
  \label{eq:comp_res}
\end{equation}
at every outer iteration $k$ and adjusts $t_k$ accordingly. For
stationarity screening only, the code also tracks the biactive
residual
\begin{equation}
  \eta_k \;=\; \max_{\ell = 1,\ldots,q} \;\min\!\bigl(|G_\ell(x^k)|,\; |H_\ell(x^k)|\bigr).
  \label{eq:biactive_res}
\end{equation}
Similar to the improvement ratio used in trust-region methods~\cite{NocedalWright2006},
we define the relative improvement
\begin{equation}
  \rho_k \;=\; \frac{r_{k-1} - r_k}{\max(r_{k-1},\, \epsilon_{\mathrm{floor}})},
  \label{eq:rho}
\end{equation}
where $\epsilon_{\mathrm{floor}}$ is a small positive guard to prevent division by zero (using the smallest normal float in the implementation). A value $\rho_k \approx 1$ indicates that the complementarity residual was nearly eliminated in one step, while small positive $\rho_k$ indicates stagnation. The update regimes are as follows.

Algorithm~\ref{alg:phase2-update} summarizes the implemented adaptive
update logic. The code realizes this logic through five branches, with
the stagnation regime split into two subcases.

\begin{algorithm}[t]
\caption{Phase~IIa Adaptive $t$-Update}
\label{alg:phase2-update}
\begin{algorithmic}[1]
\Require Current $t_k$, residuals $(r_k,r_{k-1})$, stagnation counter $s$, pathing counter $c$, biactive count history, factors $(\kappa,\alpha_L,\alpha_U)$, iteration index $k$, remaining jump budget $b$
\Ensure Next regularization parameter $t_{k+1}$, updated counters $(s,c,b)$
\State Compute relative improvement $\rho_k$ from~\eqref{eq:rho}
\State Update stagnation counter: \textbf{if} $\rho_k < 0.05$ \textbf{then} $s \gets s+1$ \textbf{else} $s \gets 0$
\State Set \textit{biactive\_stable} $\gets$ (last 3 iterates share the same $|\mathcal{I}_{00}|$)
\State Evaluate pathing condition~\eqref{eq:pathing}: \textit{cond} $\gets$ ($k \ge 4$ \textbf{and} \textit{biactive\_stable} \textbf{and} $\alpha_L \le \rho_k \le \alpha_U$)
\State Update pathing counter: \textbf{if} \textit{cond} \textbf{then} $c \gets c+1$ \textbf{else} $c \gets \max(0,c-1)$
\State Set \textit{pathing} $\gets$ ($c \ge 2$)
\If{\textit{pathing} \textbf{and} $\rho_k > 0.5$}
  \State \Comment{Path 1: superlinear mode}
  \State $t_{k+1} \gets \kappa^2\, t_k$
\ElsIf{\textit{pathing} \textbf{and} $\rho_k > 0.1$}
  \State \Comment{Path 2: fast geometric mode}
  \State $t_{k+1} \gets \min(\kappa,\,0.5)\,t_k$
\ElsIf{$s \ge 4$}
  \State \Comment{Path 3a: aggressive stagnation jump; reset $s \gets 0$}
  \State \textbf{if} $b>0$ \textbf{then} $t_{k+1} \gets \kappa^2\, t_k$, $b \gets b-1$, $s \gets 0$ \textbf{else} $t_{k+1} \gets \kappa\, t_k$
\ElsIf{$s \ge 2$}
  \State \Comment{Path 3b: post-stagnation fast}
  \State $t_{k+1} \gets \min(\kappa,\,0.5)\,t_k$
\Else
  \State \Comment{Path 4: conservative mode}
  \State $t_{k+1} \gets \kappa\, t_k$
\EndIf
  \State $t_{k+1} \gets \max(10^{-14},\, t_{k+1})$
\State \Return $t_{k+1}$
\end{algorithmic}
\end{algorithm}

\paragraph{Path 1}
The algorithm monitors whether convergence is sustained and the biactive set has stabilised. Specifically, define the \emph{pathing condition} as
\begin{equation}
  \alpha_L \;\le\; \rho_k \;\le\; \alpha_U,
  \label{eq:pathing}
\end{equation}
with $\alpha_L = 0.2$ and $\alpha_U = 5.0$, where $\rho_k$ is the relative improvement~\eqref{eq:rho}. The solver is said to be in \emph{pathing mode} at iteration $k$ when~\eqref{eq:pathing} holds, the biactive count $|\mathcal{I}_{00}|$ has been constant over the last three iterates, and $k \ge 4$. Pathing mode is \emph{sustained} when the pathing counter $c \ge 2$ (i.e., the condition has held for at least two consecutive iterations). When pathing mode is sustained and $\rho_k > 0.5$---indicating strong, steady convergence---the algorithm uses an extra-aggressive, \emph{superlinear} reduction:
\begin{equation}
  t_{k+1} = \kappa^2 \, t_k.
  \label{eq:superlinear}
\end{equation}

\paragraph{Path 2}
When pathing mode is sustained and $\rho_k > 0.1$, the algorithm maintains momentum with a capped geometric reduction:
\begin{equation}
  t_{k+1} = \min(\kappa,\,0.5)\, t_k.
  \label{eq:fast}
\end{equation}
The cap at $0.5$ prevents overly aggressive reduction when $\kappa$ is set close to~$1$, ensuring robustness across a wide range of user-supplied $\kappa$ values.

\paragraph{Path 3}
Stagnation is pathed by a counter $s$ that increments whenever $\rho_k < 0.05$ and resets to zero otherwise. Two sub-paths handle escalating stagnation. When $s \ge 4$, an aggressive jump is executed and the counter is reset:
\begin{equation}
  t_{k+1} = \kappa^2\, t_k, \qquad (s \gets 0).
  \label{eq:stagnation_jump}
\end{equation}
With the default $\kappa = 0.5$ this performs an aggressive reduction ($t_{k+1} = 0.25 t_k$), forcing the solver to move past the stagnation point. When $2 \le s < 4$, a milder post-stagnation reduction is used:
\begin{equation}
  t_{k+1} = \min(\kappa,\,0.5)\, t_k.
  \label{eq:post_stagnation}
\end{equation}

\paragraph{Path 4}
In all other cases---pathing mode is not sustained and the stagnation count is below threshold---the algorithm falls back to the standard geometric reduction:
\begin{equation}
  t_{k+1} = \kappa\, t_k.
  \label{eq:slow}
\end{equation}

An alternative to the implemented update is to set $t_{k+1} = r_k$
directly, following a complementarity-residual-aware philosophy
inspired by the theoretical sensitivities noted
in~\cite{HoheiselKanzowSchwartz2013}. This update can converge faster
but is slightly less robust on problems with oscillatory homotopy
residuals. In practice, we use the adaptive rule above as the default.

\subsubsection{Implementation safeguards for difficult iterations}
In addition to the core $t$-update, the implementation includes several
practical safeguards for difficult problems. After Phase~I, if the
warm-start is already nearly complementarity-feasible, the code can
short-circuit directly into Phase~III or fast-forward the initial
$t_k$. During Phase~II, every outer NLP
solve is protected by a wall-clock guard, and solver failures trigger
the fallback chain described in
Section~\ref{sec:solveRegNLPs} before the current iterate is abandoned.

The aggressive stagnation jump (Path~3a in Algorithm~\ref{alg:phase2-update})
is governed by a hard cap of $b = 500$ total jump triggers; exceeding
this cap prevents unbounded adaptive cycling and the outer loop
declares stagnation. Additionally, the implementation monitors an
\emph{outer stagnation guard}: if the complementarity residual changes
by less than $0.1\%$ over a sliding window of 25~iterations, the outer
loop exits early. Restoration is triggered only when the homotopy
residual exceeds $10\times\varepsilon_{\mathrm{tol}}$ relative to the
current residual (the \emph{restoration-comp factor}), and restoration
is suppressed entirely once the total restoration count exceeds
10~times the restoration cap (the \emph{high-restoration skip threshold}).

Restoration is invoked only when the sign test fails on a nonempty
biactive set and the homotopy residual is still materially above the
target tolerance. The implementation caps the total number of
restorations at 50 and monitors a stagnation window of 8~consecutive
restoration failures, after which an \emph{early BNLP rescue} is
attempted: if $r_{\mathrm{best}} \leq 1000\,\varepsilon_{\mathrm{tol}}$,
the BNLP polish (see Section~\ref{sec:phase_iii}) is invoked at
multiple active-set tolerances before declaring
\texttt{restoration\_stagnation}. If $t_k$ is clamped to the floor
$10^{-14}$ and the complementarity residual is numerically unchanged
for 20~consecutive iterations (the \emph{floor stagnation window}), the
outer loop exits early.

When Phase~II exits with $\varepsilon_{\mathrm{tol}} < r_{\mathrm{best}}
\leq 1000\,\varepsilon_{\mathrm{tol}}$, a multi-strategy refinement
attempts to push the complementarity residual below tolerance.
\emph{Gentle $t$ reduction} solves $\mathrm{NLP}(t)$ at
progressively smaller $t$ values (factors $0.5$, $0.1$, $0.01$
of the current estimate), rejecting any solution whose iterate
norm diverges by more than $10\times$ from the current point.
\emph{BNLP polish} tries the BNLP~\eqref{eq:phase_i} at
multiple active-set tolerances; accept only if it improves
$r_{\mathrm{comp}}$.
\emph{Partition flipping} flips, for each biactive index, the
active-set assignment $G_\ell = 0 \leftrightarrow H_\ell = 0$
and re-solves the BNLP to find a feasible partition.
If all strategies fail but $r_{\mathrm{best}} \leq 10\,
\varepsilon_{\mathrm{tol}}$, the point is accepted under a relaxed
tolerance and classified as C-stationary.  This three-strategy cascade
recovers problems whose homotopy residual is only slightly above the
target tolerance.


\subsubsection{Restoration phase}
\label{sec:restoration}
When the multiplier sign test fails at iteration~$k$ (i.e., some
biactive index has a sufficiently negative entry in
$\gamma^{G,k}$ or $\gamma^{H,k}$), the current iterate~$x^k$ is not
S-stationary. If the homotopy residual is still meaningfully above the
target tolerance, MPECSS starts a \emph{restoration phase} that
modifies the point and re-solves $\text{NLP}(t_k)$. In the
implementation, restoration is also used as a recovery mechanism after
recoverable solver failures (except unrecoverable statuses such as hard
invalid-number failures).

MPECSS uses three recovery strategies, tried one after another.

\paragraph{R1 Strategy}
For each biactive index $\ell \in \mathcal{I}_{00}(x^k)$, we randomly perturb the iterate along either $\nabla G_\ell$ or $\nabla H_\ell$:
\begin{equation}
  x^k_{\mathrm{pert}} = x^k + \epsilon s \frac{\nabla G_\ell(x^k)}{\|\nabla G_\ell(x^k)\|},
  \label{eq:random_perturb}
\end{equation}
where $s \sim \mathrm{Uniform}(0.5, 2.0)$ and $\epsilon$ is the magnitude. This forces one complementarity function positive, resolving the simultaneous zero. This randomized perturbation acts as a novel practical heuristic to abruptly escape non-optimal valleys before attempting continuous escape directions.

\paragraph{R2 Strategy}
This strategy calculates a deterministic escape step based on multiplier magnitude dominance:
\begin{equation}
  d_\ell = \mathrm{sign}(\gamma^{G,k}_\ell - \gamma^{H,k}_\ell).
  \label{eq:sign_direction}
\end{equation}
The iterate explores the direction $v_\ell^k$ of the larger multiplier via:
\begin{equation}
  x^k_{\mathrm{esc}} = x^k + \alpha \sum_{\ell} v_\ell^k,
  \label{eq:directional_escape}
\end{equation}
where $\alpha > 0$ halves upon consecutive failure.

\paragraph{R3 Strategy}
As a final fallback, we penalize the objective with:
\begin{equation}
  f_{\mathrm{reg}}(x) = f(x) + \gamma \sum_{\ell} \frac{(G_\ell - H_\ell)^2}{|G_\ell| + |H_\ell| + \epsilon}
  \label{eq:quad_reg}
\end{equation}
to force the solver to pick a corner. The problem is re-solved from $x^k$, doubling weight $\gamma$ on consecutive failure. This penalty approach is inspired by elastic mode methods for MPECs~\cite{Anitescu2005,HuRalph2004}.

\paragraph{Order of strategies}
The three strategies are tried in cascade order: R1 $\to$ R2 $\to$ R3. The first strategy that produces a successful NLP re-solve is accepted immediately, and its result is returned to the main loop (Algorithm~\ref{alg:phase2-main}). The sign test is then re-evaluated by the main loop on the returned point; if it passes, the restored iterate replaces $x^k$ and the algorithm continues from this improved starting point. The R3 strategy internally attempts up to three re-solves with increasing penalty weight $\gamma$ (doubling $\gamma$ on each failure) before reporting failure. If all three strategies fail to produce a feasible re-solve, the restoration returns failure and the main loop continues with the standard $t$-update (Section~\ref{sec:adaptive_t}).

Algorithm~\ref{alg:phase2-restoration} gives the restoration workflow used when the sign test fails or recoverable solver failure is detected.

\begin{algorithm}[t]
\caption{Phase~IIb Restoration Subroutine at Fixed $t_k$}
\label{alg:phase2-restoration}
\begin{algorithmic}[1]
\Require Current iterate $x^k$, multipliers $(\gamma^{G,k},\gamma^{H,k})$, biactive set $\mathcal{I}_{00}(x^k)$
\Ensure Restored iterate (caller re-evaluates sign test) or \textbf{nil}
\If{$\mathcal{I}_{00}(x^k) = \emptyset$}
  \State \Return \textbf{nil}
\EndIf
\For{strategy $s$ in $\{\mathrm{R1},\,\mathrm{R2},\,\mathrm{R3}\}$}
  \If{$s = \mathrm{R1}$}
    \State Build random Jacobian-guided perturbation~\eqref{eq:random_perturb} $\to$ $x^k_{\mathrm{trial}}$
    \State Re-solve NLP$(t_k)$ from $x^k_{\mathrm{trial}}$
  \ElsIf{$s = \mathrm{R2}$}
    \State Build directional escape step~\eqref{eq:sign_direction}--\eqref{eq:directional_escape} $\to$ $x^k_{\mathrm{trial}}$
    \State Re-solve NLP$(t_k)$ from $x^k_{\mathrm{trial}}$
  \Else
    \State Solve penalised subproblem~\eqref{eq:quad_reg} (up to 3 attempts, doubling $\gamma$ on each failure)
  \EndIf
  \If{re-solve succeeds}
    \State \Return restored iterate (sign test checked by caller)
  \EndIf
\EndFor
\State \Return \textbf{nil} (caller continues with standard $t$-update)
\end{algorithmic}
\end{algorithm}

The restoration phase is a practical method that does not change the theoretical convergence guarantees of Scholtes's regularization (Theorem~\ref{thm:scholtes}). When recovery succeeds, the algorithm follows a regularization path toward a stronger stationarity point. When it fails, the algorithm returns to the standard parameter reduction, still ensuring convergence to at least a C-stationary point under MPEC-LICQ.


\subsection{Phase III: B-Stationarity Certification}
\label{sec:phase_iii}
After Phase~II terminates with a candidate $x^*$, MPECSS enters a
certification and refinement stage inspired by MPECopt-style active-set
refinement and directional certification~\cite{NurkanovicLeyffer2025}.
In the core solver, this stage first performs the LICQ/LPEC logic on
the best complementarity-feasible point. In the benchmark pipeline used
for the numerical experiments, it is then followed by BNLP polishing, an
LPEC refinement loop, and a final post-check for B-stationarity.

Algorithm~\ref{alg:phase3} summarizes the reported certification and
refinement sequence.

\begin{algorithm}[t]
\caption{Phase~III B-Stationarity Certification and Refinement}
\label{alg:phase3}
\begin{algorithmic}[1]
\Require Candidate point $x^*$ from Phase~II
\Ensure Final status and refined point
\If{$x^*$ is complementarity-feasible}
  \State Check MPEC-LICQ at $x^*$
  \If{sign test passes \textbf{and} MPEC-LICQ holds}
    \State Certify $x^*$ as B-stationary by S/B equivalence
  \Else
    \State Run LPEC-based certification on $x^*$
  \EndIf
\EndIf
\State Run BNLP polishing; accept only if complementarity remains small and the objective does not worsen materially
\State Run trust-region LPEC refinement and a final B-stationarity post-check
\State Return the final point and certification status
\end{algorithmic}
\end{algorithm}

\emph{Direct certification under sign test plus MPEC-LICQ:} If the final iterate passes the multiplier sign test and MPEC-LICQ holds at that point, then the implementation certifies B-stationarity by the equivalence between S- and B-stationarity under MPEC-LICQ.

\emph{BNLP polishing:} Even a good solution can be numerically rough. We solve a final simplified problem (BNLP) by fixing an active-set partition. Given index sets $\mathcal{I}_1 = \{\ell : G_\ell(x^*) \approx 0\}$ and $\mathcal{I}_2 = \{\ell : H_\ell(x^*) \approx 0\}$, the BNLP subproblem is:
\begin{equation}
\begin{aligned}
  \text{BNLP}(x^*,\mathcal{I}_1,\mathcal{I}_2): \quad & \min_{x \in \mathbb{R}^n} \; f(x) \\
  & \text{s.t.} \quad g(x) \leq 0, \; h(x) = 0, \\
  & \quad\; G_\ell(x) = 0 \; (\ell \in \mathcal{I}_1), \; H_\ell(x) \geq 0 \; (\ell \in \mathcal{I}_1), \\
  & \quad\; H_\ell(x) = 0 \; (\ell \in \mathcal{I}_2), \; G_\ell(x) \geq 0 \; (\ell \in \mathcal{I}_2).
\end{aligned}
\label{eq:bnlp}
\end{equation}
For biactive indices ($\ell \in \mathcal{I}_{00}$), either $G_\ell$ or
$H_\ell$ must be assigned to the equality branch; the implementation
first uses the primary partition and then tries limited single-flip
alternatives. In the core solver, a polished point is accepted only if
its complementarity residual remains within the configured tolerance and
its objective value does not worsen by more than a small
relative/absolute tolerance. In the benchmark harness, an additional
post-check may still relabel near-feasible failures with
$r_{\mathrm{comp}} \le 10\,\varepsilon_{\mathrm{tol}}$ after the final
B-stationarity check.

\emph{LPEC-based certification and refinement:} If sign test plus LICQ
do not already certify the point, we use the geometric LPEC check to
determine whether a feasible descent direction exists~\cite{LuoPangRalph1996,FletcherLeyfferRalphScholtes2006}.
The implemented LP model includes the active original constraints,
active complementarity-side equalities, bound-consistent trust-region
limits $|d_i| \le \rho$, and branch-dependent inequalities for both
lower-bound and upper-bound biactive complementarity indices. For a
point with $|\mathcal{I}_{00}|$ biactive pairs, the implementation
enumerates all XOR branches (either the $G_\ell$ side or the $H_\ell$
side stays active per biactive index), capped at $2^{15}$ branches for
tractability, and solves the resulting LPs with SciPy's
\texttt{linprog(..., method='highs')} backend~\cite{HuangfuHall2018}.
If the branch enumeration times out or hits the branch cap without
finding a descent direction, the outcome is treated as
\texttt{stationarity\_unverifiable} rather than as a certificate. In
benchmark mode, this certification is followed by a trust-region
refinement loop with parameters
$\rho_{\mathrm{init}} = 10^{-2}$, $\gamma_L = 0.1$,
$N_{\mathrm{out}} = 30$, $N_{\mathrm{in}} = 15$,
$tol_B = 10^{-10}$, and $tol_{\mathrm{comp}} = 10^{-8}$.

In the reported experiments, every problem is post-processed by the full
benchmark pipeline: the Phase~II output is passed through BNLP
polishing, LPEC refinement, and a final B-stationarity post-check,
even when the initial sign test is favorable. This benchmark harness is
therefore slightly richer than a bare call to \texttt{run\_mpecss()},
and the numerical section reports the output of that harness rather than
the raw Phase~II endpoint alone.


% =========================================================================
