\section{Convergence Analysis}
\label{sec:convergence}

This section separates two distinct claims that are both relevant for
MPECSS. First, the Phase~II regularization path inherits the classical
Scholtes regularization guarantee, under standard assumptions, converging to a C-stationary point. Second, the
implementation applies a finite-iterate certification layer in
Phase~III, which may upgrade the returned candidate to a certified
B-stationary point when the LICQ or LPEC checks succeed.

\subsection{Assumptions and Convergence Framework}
\label{sec:assumptions}

Throughout this section, we make the following standard assumptions, which are common in the MPEC literature~\cite{LuoPangRalph1996,ScheelScholtes2000,Scholtes2001}:

\begin{assumption}[Smoothness and Boundedness]
\label{ass:smoothness}
The functions $f$, $g_i$ ($i=1,\ldots,m$), $h_j$ ($j=1,\ldots,p$), $G_\ell$, and $H_\ell$ ($\ell=1,\ldots,q$) are twice continuously differentiable. The feasible set
\[
  \mathcal{F} = \{x \in \mathbb{R}^n : g_i(x) \leq 0, h_j(x) = 0, G_\ell(x) \geq 0, H_\ell(x) \geq 0, G_\ell(x) H_\ell(x) = 0\}
\]
is non-empty and bounded.
\end{assumption}

\begin{assumption}[Constraint Qualification]
\label{ass:cq}
Whenever a limit point or returned candidate $x^*$ is discussed, the
specific MPEC constraint qualification required by the included result is
assumed to hold at $x^*$. In particular, MPEC-LICQ
(Definition~\ref{def:mpec_licq}) is used in the direct asymptotic
C-stationarity result below and in the Phase~III S/B-equivalence
shortcut, while MPEC-MFCQ (Definition~\ref{def:mpec_mfcq}) is relevant
only for weaker multiplier-existence statements.
\end{assumption}

\begin{assumption}[NLP Solver Accuracy]
\label{ass:nlp_accuracy}
At each outer iteration $k$, the computed point $x^k$ is a sufficiently
accurate approximate KKT point of the regularized subproblem
$\text{NLP}(t_k)$, and its primal feasibility violations are at the
level of numerical solver tolerances. The analysis below therefore
requires only sufficiently accurate inexact subproblem solves; it does
not rely on an explicit schedule of the form $\varepsilon_k = O(t_k)$.
\end{assumption}

% Assumption~\ref{ass:nlp_accuracy} matches the released code, where
% IPOPT~\citep{WachterBiegler2006} is the default NLP engine and the
% benchmark configuration uses fixed inner tolerances together with outer
% stopping tests and Phase~III post-processing. This is consistent with
% \cite{KanzowSchwartz2015}, which shows that sufficiently accurate
% inexact solves preserve the same convergence type as the exact Scholtes
% scheme.

\subsection{Convergence of the Regularization Sequence}
\label{sec:conv_reg_seq}

The foundation of MPECSS is Scholtes's regularization~\cite{Scholtes2001}, which has been rigorously studied in the
literature~\cite{HoheiselKanzowSchwartz2013,RalphWright2004,SteffensenUlbrich2010}. The following theorem summarizes the exact Scholtes result together with the practical variant for approximate solves.

\begin{theorem}[Scholtes's Regularization Convergence~\citep{Scholtes2001,KanzowSchwartz2015}]
\label{thm:scholtes}
Let $\{x^k\}$ be a bounded sequence of exact KKT points of the
regularized subproblems $\text{NLP}(t_k)$ with $t_k > 0$ and
$t_k \to 0$. Then every limit point $x^*$ of $\{x^k\}$ is
feasible for the original MPEC~\eqref{eq:general_mpec}. If, in
addition, MPEC-LICQ (Definition~\ref{def:mpec_licq}) holds at $x^*$,
then $x^*$ is C-stationary (Definition~\ref{def:mpec_stationarity}).
Moreover, sufficiently accurate approximate KKT sequences for the
Scholtes subproblems have the same limiting convergence type.
\end{theorem}

\begin{proof}
This is the classical Scholtes convergence result, while the statement
for sufficiently accurate approximate subproblem solves is the corresponding
extension given in~\cite{KanzowSchwartz2015}. As $t_k \to 0$, the
relaxed product constraints force complementarity in the limit, hence
every limit point is MPEC-feasible. Passing to the limit in the
subproblem KKT systems then achieves C-stationarity under MPEC-LICQ; see
\cite{Scholtes2001,KanzowSchwartz2015} for full details.
\end{proof}

Theorem~\ref{thm:scholtes} guarantees only C-stationarity, which is
strictly weaker than B-stationarity in the
hierarchy~\eqref{eq:stat_hierarchy}. Standard Scholtes regularization
does not have the multiplier sign conditions associated
with S-stationarity or B-stationarity.

\begin{lemma}[Adaptive $t$ Maintains the Scholtes Guarantee]
\label{lem:adaptive_t_convergence}
Under Assumptions~\ref{ass:smoothness}--\ref{ass:nlp_accuracy}, if the
sequence $\{t_k\}$ generated by the implemented adaptive rule remains
positive and satisfies $t_k \to 0$, then the adaptive rule does not
change the classical Scholtes regularization's conclusion; any limit point
of a bounded Phase~II sequence of approximate KKT iterates is
MPEC-feasible and C-stationary. Any stronger B-stationarity statement
comes from the separate finite-iterate Phase~III certification stage.
\end{lemma}

The multiplier sign test in MPECSS is an implementation-level screening
device rather than a separate convergence theorem. At each
outer iteration $k$, after solving $\text{NLP}(t_k)$, the algorithm
extracts the converted multipliers $(\gamma^{G,k}, \gamma^{H,k})$ and
checks the sign conditions at biactive indices
$\ell \in \mathcal{I}_{00}(x^k)$:
\begin{equation}
  \gamma^{G,k}_\ell \geq -\tau, \quad \gamma^{H,k}_\ell \geq -\tau,
  \label{eq:sign_test_conditions}
\end{equation}
with the fixed sign-test tolerance $\tau$ (default $10^{-6}$). If all biactive indices satisfy
these conditions, the iterate $x^k$ is accepted as an
S-stationarity candidate.

\begin{theorem}[S-stationarity Implies B-stationarity {\cite{ScheelScholtes2000}}]
\label{thm:s_implies_b}
Let $x^*$ be feasible for~\eqref{eq:general_mpec}. If $x^*$ is S-stationary (Definition~\ref{def:mpec_stationarity}), then $x^*$ is B-stationary (Definition~\ref{def:mpec_stationarity}). This implication holds without requiring any constraint qualification.
\end{theorem}

In practice, the sign test is a fast algebraic check performed after
each NLP solve. The biactive indices are detected adaptively with a
tolerance $\sigma_k = \max(10^{-8}, \sqrt{t_k})$, ensuring that the
screening set scales with the regularization parameter even though the
sign threshold $\tau$ itself is fixed. The sign test is a quick check
that a point may be S-stationary, but it does not prove it. Finite
solver tolerances and possible multiplier nonuniqueness can still
affect the result, so Phase~III is used to verify the point.

\subsection{Role of Feasibility Phase}
\label{sec:restoration_convergence}

Phase~I is a preprocessing heuristic rather than part of the standard
Scholtes's regularization convergence proof. As described in
Section~\ref{sec:feasibility}, the implementation cycles through
several smooth feasibility models of the generic
form~\eqref{eq:phase_i}. The first regime solves
\begin{equation}
  \text{Phase I:} \quad \min_{x} \sum_{\ell=1}^{q} (G_\ell(x) H_\ell(x))^2 \quad \text{s.t.} \quad g_i(x) \leq 0, h_j(x) = 0, G_\ell(x) \geq 0, H_\ell(x) \geq 0.
  \label{eq:phase_i_revisit}
\end{equation}
and subsequent fallback regimes replace only the objective while keeping
the same non-complementarity constraints. A successful Phase~I run can
substantially reduce the complementarity residual and provide a better
warm start for Phase~II, but the core convergence statement
of Section~\ref{sec:conv_reg_seq} does not depend on a separate
Phase~I theorem.

\begin{lemma}[Phase I Provides a Useful Warm Start]
\label{lem:phase_i_warm_start}
Suppose the Phase~I procedure returns a point $\bar{x}$ satisfying the
original inequality, equality, and bound constraints to numerical
precision and having small complementarity residual. Then $\bar{x}$ is
feasible for the non-complementarity constraints and approximately
complementarity-feasible for the MPEC. When Phase~II is initialized with
such a point, the early sign-test logic and the fast-forward heuristic
described in Section~\ref{sec:feasibility} may shorten the homotopy path
substantially. For large problems, however, the implementation may skip
Phase~I entirely, so the convergence proof depends on the Phase~II Scholtes sequence.
\end{lemma}




\subsection{B-Stationarity Certification and LPEC Post-Solve}
\label{sec:bstat_certification}

After Phase~II returns a complementarity-feasible point, MPECSS
executes a \emph{B-stationarity certification phase} (Phase~III) to
verify whether the returned finite iterate is B-stationary.

When the sign test passes at the final point $x^*$ and MPEC-LICQ
holds at $x^*$, the implementation certifies $x^*$ as B-stationary by
the equivalence between S-stationary and B-stationary under LICQ. If MPEC-LICQ
does not hold, or if the sign-test shortcut is unavailable, MPECSS
solves an LPEC check to provide a fall-back
geometric verification:

\begin{definition}[ LPEC Check~\citep{LuoPangRalph1996,FletcherLeyfferRalphScholtes2006,NurkanovicLeyffer2025}]
Given a point $x^*$, the directional LP solved inside the
implementation has the form
\begin{equation}
\begin{aligned}
  \text{LP}(x^*,\rho,\beta): \quad & \min_{d \in \mathbb{R}^n} \quad && \nabla f(x^*)^\top d \\
  & \text{s.t.} \quad && \nabla g_i(x^*)^\top d \leq 0 \quad (i \in \mathcal{I}_g(x^*)), \\
  & && \nabla h_j(x^*)^\top d = 0 \quad (j = 1, \ldots, p), \\
  & && \nabla G_\ell(x^*)^\top d = 0 \quad (\ell \in \mathcal{I}_G), \\
  & && \nabla H_\ell(x^*)^\top d = 0 \quad (\ell \in \mathcal{I}_H), \\
  & && \text{branch constraints determined by } \beta \text{ on each biactive index}, \\
  & && -\rho \le d_i \le \rho,\qquad x^* + d \in [x_L, x_U].
\end{aligned}
\label{eq:lpec}
\end{equation}
Here $\beta$ indexes the XOR branch choice for each biactive pair:
either the $G$-side or the $H$-side remains active, with distinct
inequality orientation for lower-bound and upper-bound
box-complementarity cases. The implementation enumerates these branches
up to a hard cap of $2^{15}$ and solves each LP with a finite wall-clock
timeout. A completed enumeration with objective value $\ge 0$ certifies
that no descent direction was found within this implemented model; a
timeout or cap-exceeded run achieves an \emph{unverifiable} outcome rather
than a certificate.
\end{definition}

\begin{proposition}[LPEC Certification of B-Stationarity]
\label{thm:lpec_bstat}
If $x^*$ is feasible for~\eqref{eq:general_mpec}, the full branch
enumeration associated with~\eqref{eq:lpec} completes, and every branch
LP has optimal value $\geq 0$, then the implemented verification certifies
$x^*$ as B-stationary. Moreover, if MPEC-LICQ holds at $x^*$,
B-stationarity is equivalent to S-stationarity
(Definition~\ref{def:mpec_stationarity}). If the enumeration times out
or exceeds the branch cap before completion, the implementation does not
claim B-stationarity.
\end{proposition}


The code distinguishes between internal checking states and final
reported outcomes. Internally, a sign-test pass marks an
S-stationarity candidate. The final reported outcomes are:
\begin{itemize}
  \item[-] \emph{B-stationary (sign test + MPEC-LICQ):} The final candidate
  passes the multiplier sign test, MPEC-LICQ is verified, and the
  implementation certifies B-stationarity through S/B equivalence under
  LICQ.
  \item[-] \emph{B-stationary (LPEC certified):} Full LPEC certification
  (Proposition~\ref{thm:lpec_bstat}) completes and confirms that no
  feasible first-order descent direction exists.
  \item[-] \emph{C-stationary:} The candidate is complementarity-feasible,
  but either the sign-test shortcut fails or the completed LPEC verification
  finds a feasible descent direction, so B-stationarity cannot be
  claimed.
  \item[-] \emph{Stationarity unverifiable / non-converged:} If the LPEC
  wall-clock timeout is reached, exceeds the branch cap, or the main solve never
  reaches complementarity feasibility, the implementation reports an
  unverifiable or failed status rather than a theoretical stationarity
  label.
\end{itemize}

\subsection{Convergence Summary and Computational Implications}
\label{sec:convergence_summary}

\begin{theorem}[Convergence of MPECSS]
\label{thm:mpecss_convergence}
Under Assumptions~\ref{ass:smoothness}--\ref{ass:nlp_accuracy}, suppose
the Phase~II regularization parameters remain positive and satisfy
$t_k \to 0$. Then the following conclusions hold:
\begin{enumerate}
  \item Every limit point of a bounded subsequence of Phase~II
  approximate KKT iterates is feasible for the original
  MPEC~\eqref{eq:general_mpec}.
  \item If MPEC-LICQ holds at such an limit point, then that
  point is C-stationary by Theorem~\ref{thm:scholtes}.
  \item A finite-iterate sign-test pass is treated only as an
  S-stationarity point under the implemented tolerances; by itself,
  it is not the final reported stationarity label.
  \item If the returned point passes the sign test and MPEC-LICQ
  holds at that point, then Phase~III certifies B-stationarity by S/B
  equivalence.
  \item If the full LPEC certification completes and every branch LP has
  optimal value at least zero, then the returned point is certified
  as B-stationary.
  \item If the completed LPEC verification finds a feasible descent
  direction, then the returned point is not B-stationary and the
  implementation classifies it at most as C-stationary.
  \item If the LPEC certification times out or the branch cap is reached
  before completion and no descent direction has been proved, then the
  implementation reports \texttt{stationarity\_unverifiable} rather
  than a B-stationarity certificate.
  \item The optional benchmark-side post-processing of near-feasible
  non-converged outcomes with
  $r_{\mathrm{comp}} \le 10\,\varepsilon_{\mathrm{tol}}$ is a salvage
  heuristic, not a new convergence theorem.
\end{enumerate}
\end{theorem}


While Theorem~\ref{thm:mpecss_convergence} establishes the theoretical convergence guarantees, the practical rate still
depends strongly on problem conditioning, the chosen NLP solver, and the
effectiveness of the adaptive $t$-update.

\subsection{Local Optimality and Constraint Qualifications}
\label{sec:local_optimality}

The results in this section address first-order stationarity of feasible
points, not direct proofs of local optimality for the original
objective. Since B-stationarity is a necessary (but not sufficient)
condition for local optimality~\citep{LuoPangRalph1996,ScheelScholtes2000},
MPECSS certifies only that no first-order feasible descent direction has
been found in the Bouligand sense. Additional second-order conditions
are needed to upgrade this to local optimality; see
\cite{OutrataKocvaraZowe1998} for details.

% \text{Remark on Unbounded Feasible Sets:}
% If the feasible set is unbounded, Assumption~\ref{ass:smoothness}
% should be replaced by standard level-boundedness or coercivity
% hypotheses that guarantee the existence of bounded subsequences. Under
% those modifications, the conclusions above apply to any limit point that exists.
